On the formation of H$\alpha$ line emission around classical T Tauri stars

### 2006-04-05 TTau_Halpha/revision01/kurosawa.tex
Revised version after the refree's report
git-svn-id: https://repository.astro.ex.ac.uk/torus/trunk@693 25464f21-6099-402c-b30a-640f1d40d38b

In the currently favoured model of accretion in CTTS, the accretion
disc is disrupted by the magnetosphere, which channels the gas from
the disc onto the stellar surface (e.g., \citealt{uchida:1985};
\citealt{koenigl:1991}; \citealt{cameron:1993};
\citealt{shu:1994}). This picture is supported by recent measurements
of strong ($\sim 10^{3}\mathrm{G}$) magnetic fields in CTTS (e.g.,
\citealt{johns-krull:1999}; \citealt{symington:2005b}) and by
radiative transfer models which reproduce the gross characteristics of
observed profiles for some CTTS (\citealt*{muzerolle:2001}). In
particular the magnetospheric accretion (MA) model explains blue-ward
asymmetric emission line profiles as resulting from the partial
occultation of the flow by the stellar photosphere, while inverse
P~Cygni profiles in the MA model result from inflowing material at
near free-fall velocities seen projected against hotspots on the
stellar surface.

@@

An additional continuum source is considered for models which include 
magnetospheric accretion: as the infalling gas approaches the stellar
surface, it decelerates in a strong shock, and is heated to $\sim
10^{6}\,\mathrm{K}$.  The X-ray radiation produced in the shock will
be absorbed by the gas locally, and re-emitted  optical and UV
radiation (\citealt{koenigl:1991}; \citealt*{hartmann:1994}). This
will create hot rings where the magnetic field intersects with the
surface. We assume that the free-falling kinetic energy is thermalized
in the radiating layer, and is re-emitted as blackbody radiation with
a single temperature. With the parameters of the magnetosphere and the
star given above (Table~\ref{tab:std_parameters}), about $8$~per~cent
of the surface is covered by the hot rings. If the mass-accretion rate
is $10^{-7}\,\mathrm{M_{\sun}\, yr^{-1}}$, the ratio of this accretion
luminosity to the photospheric luminosity is about $0.5$, and the
corresponding temperature of the hot rings is about
$6400\,\mathrm{K}$.  The continuum emission from the hot rings is
taken into account when computing the line profiles.

@@

\caption{Summary of the reference classical T~Tauri star model parameters.}
\label{tab:std_parameters}

@@

We adopt the MA flow model of \citet{hartmann:1994},
as done  by \citet{muzerolle:2001} and by \citet*{symington:2005},
in which the gas accretion on to the stellar surface from the innermost
part of the accretion disc occurs through a dipolar stellar magnetic
field. The magnetic field is assumed to be so strong that the gas
flow does not affect the underlying magnetic field itself. As shown
in Fig.~\ref{fig:config_discwind} (region~2), the innermost radius ($R_{\mathrm{mi}}$)
of the magnetosphere at the equatorial plane ($z=0$) is assigned
to be same as the inner radius ($R_{\mathrm{di}}$) of the accretion
disc where the flow is truncated. In our models, $R_{\mathrm{mi}}$
and the outer radius ($R_{\mathrm{mo}}$) of the magnetosphere (at
the equatorial plane) are set to be $2.2\, \mathrm{R_{\sun}}$ and $3.0\, \mathrm{R_{\sun}}$
respectively. The geometry of the magnetic field/stream
lines is fixed for all calculations. We note that this magnetospheric geometry
is identical to the {}``small/wide'' model of \citet{muzerolle:2001}.

The magnetic field and the gas stream lines are assumed to have the
following simple form 
\begin{equation}
      r=R_{\mathrm{m}}\sin^{2}\theta
  \label{eq:dipole}
\end{equation}
\citep*[see][]{ghosh:1977} where $r$, and $\theta$ are the radial and
the polar component of a field point position vector in the accretion
stream.  $R_{\mathrm{m}}$ is the radial distance to the field
line at the equatorial plane ($\theta=\pi/2$), and its value is
restricted between $R_{\mathrm{mi}}$ and 
$R_{\mathrm{mo}}$ (Fig.~\ref{fig:config_discwind}). 
Using the field geometry above and conservation of
energy, the velocity and the density of the accreting gas along the
steam line are found as in \citet{hartmann:1994}.

The temperature structure of the magnetospheric used by
\citet{hartmann:1994} is adopted here. They computed the temperature,
assuming a volumetric heating rate which is proportional to $r^{-3}$,
by solving the energy balance of the radiative cooling rate of
\citet{hartmann:1982} and the heating rate \citep{hartmann:1994}.
\citet{martin:1996} presented a self-consistent determination of the
thermal structure of the inflowing gas along the dipole magnetic field
(equation~\ref{eq:dipole}) by solving the heat equation coupled to the
rate equations for hydrogen. He found that main heat source is
adiabatic compression due to the converging nature of the flow, and
the major contributors to the cooling process are bremsstrahlung
radiation and line emission from Ca~II and Mg~II ions. The results of
\citet{martin:1996} qualitatively agree with those of
\citet{hartmann:1994}.

@@

The local mass-loss rate per unit area ($\dot{m}$) is assumed to
be proportional to the mid-plane temperature of the disc, and is a
function of the cylindrical radius $w=\left(x^{2}+y^{2}\right)^{1/2}$, i.e.
\begin{equation}
  \dot{m}\left(w\right) \propto
       \left[ T\left(w\right) \right]  ^{4\alpha}\,\,.\label{eq:discwind-massloss-temp}
\end{equation}
The mid-plane temperature of the disc is assumed to be expressed
as a power-law in $w$; thus, $T\propto w^{q}$. Using this in the
relation above, one finds 
\begin{equation}
  \dot{m}\left(w\right)\propto w^{p}
  \label{eq:discwind-massloss-w}
\end{equation}
where $p=4\alpha\times q$. The index of the mid-plane temperature power
law is adopted from the dust radiative transfer model of
\citet{whitney:2003a} who found the innermost part of the accretion
disc has $q=-1.15$.  In order to be consistent with the collimated
disc-wind model of \citet{krasnopolsky:2003} who used $p=-5/2$, the
value of $\alpha$ is set to 0.76. The constant of proportionality in
equation~\ref{eq:discwind-massloss-w} is found by integrating
$\dot{m}$ from $R_{\mathrm{wi}}$ to $R_{\mathrm{wo}}$, and 
normalising the value to the total mass-loss rate
$\dot{M}_{\mathrm{wind}}$.

The azimuthal/rotational component of the wind velocity
$v_{\phi}\left(w,z\right)$ is computed from the Keplerian rotational
velocity at the emerging point of the stream line
i.e. $v_{\phi}\left(w_{i},0\right)=\left(GM_{*}/w_{i}\right)^{1/2}$
where $w_{i}$ is the distance from the rotational axis ($z$) to the
emerging point on the disc, and by assuming the conservation of the
specific angular momentum along a stream line:
\begin{equation}
  v_{\phi}\left(w,z\right)\, = 
  v_{\phi}\left(w_{i},0\right)\,\left(\frac{w_{i}}{w}\right)\,\,.
  \label{eq:discwind-toroidal-velocity}
\end{equation}
Based on the canonical $\beta$ velocity law of hot stellar winds
(c.f.~\citealt*{castor:1975}), the poloidal component of the wind
velocity ($v_{p}$) parameterised as:
\begin{equation}
  v_{p}\left(w_{i},l\right) = c_{\mathrm{s}}\left(w_{i}\right) +
  \left[f\, v_{\mathrm{esc}}-c_{\mathrm{s}}\left(w_{i}\right)\right]\left(1-\frac{R_{s}}{l+R_{s}}\right)^{\beta}
  \label{eq:discwind-poloidal-velocity}
\end{equation}  
where $c_{\mathrm{s}}$, $f$, and $l$ are the sound speed at the wind
launching point on the disc, the constant scale factor of the
asymptotic terminal velocity to the local escape velocity (from the
wind emerging point on the disc), and the distance from the disc
surface along stream lines respectively. $R_{s}$ is the wind scale
length, and its value is set to 10~$R_{\mathrm{mi}}$ by following
\citet{long:2002}.

@@

  \caption{The dependency of the disc-wind density and velocity on the wind
    acceleration parameter $\beta$. The wind density $\rho$ (right panel)
    and the poloidal velocity component $v_{p}$ (left panel) along the
    stream line starting from the mid point of the wind launching zone,
    i.e. $\left ( w,z \right ) = \left ( w_{\mathrm{mid}}, 0 \right )$
    where $w_{\mathrm{mid}} = \left( R_{\mathrm{wi}}+R_{\mathrm{wo}}
    \right ) / 2$, are shown as a function of the distance ($l$) from the
    wind launching point (c.f.~
    equations~\ref{eq:discwind-poloidal-velocity} and
    \ref{eq:discwind-density}). The azimuthal velocity component
    ($v_{\phi}$), which is independent of $\beta$
    (c.f. equation~\ref{eq:discwind-toroidal-velocity}), is also shown in
    the right panel for comparison. The density is normalised to the
    density $\rho_{0}$ at the wind launching point for the $\beta = 1.0$ case.
    The $v_{p}$ reaches the terminal velocity by 100~au for all
    $\beta$. In the far field ($l>10$~au), the density is approximately
    proportional to $\sim l^2$. }

@@

\subsubsection{Dust model}

@@

We further modified {\sc torus} by replacing the Monte Carlo line
transfer algorithm with a direct integration method
(c.f. \citealt{mihalas:1978}) for computing the observed flux as a
function of frequence.  The integration of the flux is performed in
the cylindrical coordinate system $\left(p,\, q,\, t\right)$ which is
obtained by rotating the original stellar coordinate system
$\left(\rho,\,\phi,\, z\right)$ around the $y$ axis by the inclination
of the line of sight.  Note that the $t$-axis coincides with the line
of sight with this rotation. The observed flux ($F_{\nu}$) is given
by:
\begin{equation}
  F_{\nu}=\frac{1}{4\pi d^{2}}
  \int_{0}^{p_{\mathrm{max}}}\int_{0}^{2\pi}p\,\sin q\,
  I_{\nu}\mathrm{\, d}q\,\mathrm{d}p
  \label{eq:flux_integral}
\end{equation}
where $d$, $p_{\mathrm{max}}$, and $I_{\nu}$ are the distance to
an observer, the maximum extent to the model space in the projected
(rotated) plane, and the specific intensity ($I_{\nu}$) in the direction
on observer at the outer boundary. For a given ray along $t$, the
specific intensity is given by:
\begin{equation}
       I_{\nu} = I_{0}e^{-\tau_{\infty}} + 
           \int_{t_{0}}^{t_{\infty}}S_{\nu}\left(t\right)\,
                e^{-\tau}\mathrm{d}\tau
 \label{eq:formal_sol_integral}
\end{equation}
where $I_{0}$ and $S_{\nu}$ are the intensity at the boundary on the
opposite to the observer and the source function
($\eta_{\nu}/\chi_{\nu}$) of the stellar atmosphere/wind at a
frequency $\nu$. For a ray which intersects with the stellar core,
$I_{0}$ is computed from the stellar atmosphere model of
\citet{kurucz:1979} as described in
Section~\ref{sub:Continuum-Source}, and $I_{0}=0$ otherwise. If the
the ray intersects with the hot ring on the stellar surface created by
the accretion stream, we set
$I_{0}=B_{\nu}\left(T_{\mathrm{ring}}\right)$ where $B_{\nu}$ is the
Planck function and $T_{\mathrm{ring}}$ is the temperature of the hot
ring.  The initial position of each ray is assigned to be at the
centre of the surface element ($\mathrm{d}A=p\,\sin
q\,\mathrm{d}q\,\mathrm{dp}$).  The code execution time is
proportional to $n_{p}\, n_{q}\, n_{\nu}$ where $n_{p}$ and $n_{q}$
are the number of cylindrical radial and angular points for the flux
integration, and $n_{\nu}$ is the number of frequency points. In the
models presented in the following section, $n_{p}=180$, $n_{q}=100$,
and $n_{\nu}=101$ are used unless specified otherwise. A linearly
spaced radial grid is used for the area where the ray intersects with
magnetosphere, and a logarithmically spaced grid is used for the wind
and the accretion disc regions.

@@

For a point in the magnetosphere and the wind flows, the emissivity
and the opacity of the media are given as:
\begin{equation}
  \left\{ 
  \begin{array}{rcl}
    \eta_{\nu} & = & \eta_{\mathrm{c}}^{\mathrm{H}}+\eta_{l}^{\mathrm{H}}\\
    \chi_{\nu} & = & \chi_{c}^{\mathrm{H}}+\chi_{l}^{\mathrm{H}}
    +\sigma_{\mathrm{es}}
  \end{array}
  \right.
  \label{eq:emiss_opa_gas}
\end{equation}
where $\eta_{\mathrm{c}}^{\mathrm{H}}$ and $\eta_{l}^{\mathrm{H}}$ are
the continuum and line emissivity of
hydrogen. $\chi_{c}^{\mathrm{H}}$, $\chi_{l}^{\mathrm{H}}$, and
$\sigma_{\mathrm{es}}$ are the continuum, line opacity
(equation~\ref{eq:line_opacity}) of hydrogen, and the electron
scattering opacity. Similarly, for a point in the accretion disc,
\begin{equation}
  \left\{ 
  \begin{array}{rcl}
    \eta_{\nu} & = & 0\\
    \chi_{\nu} & = & \kappa_{\mathrm{abs}}^{\mathrm{dust}}
    +\kappa_{\mathrm{sca}}^{\mathrm{dust}}
  \end{array}
  \right.
  \label{eq:emiss_opa_dust}
\end{equation}
where $\kappa_{\mathrm{abs}}^{\mathrm{dust}}$ and
$\kappa_{\mathrm{sca}}^{\mathrm{dust}}$ are the dust absorption, and
scattering opacity which are calculated using the dust property
described in Section~\ref{eq:grain-dist-function}.  We neglect the
emissivity from the disc: since the disc masses of CTTS are rather small
($\sim 0.01\,\mathrm{M_{\sun}}$) and low temperature ($\lesssim 
1600\,\mathrm{K}$), the continuum flux contribution at H$\alpha$
wavelength is expected to be negligible (e.g.~\citealt{chiang:1997}).

@@

Using the reference model parameters (Table~\ref{tab:std_parameters}) for the
central star and the magnetosphere, we examine the dependency of
H$\alpha$ on the temperature ($T_{\mathrm{max}}$) of the accretion flow
and the mass accretion rate ($\dot{M}_{\mathrm{acc}}$), as
done by \citet{muzerolle:2001} for H$\beta$. We note that here we have
 computed the hot ring temperature self-consistently, whereas \citet{muzerolle:2001} used
a constant hot ring temperature ($8000\,\mathrm{K}$) for most of
their models.  The accretion luminosity ($L_{\mathrm{acc}}$) for models
with $\dot{M}_{\mathrm{acc}}=10^{-7}\,\mathrm{M_{\sun}\, yr^{-1}}$ is
about a half of the photospheric luminosity.

@@

The dependency of the line profile on inclination angle ($i$) is
demonstrated in Fig.~\ref{fig:acc_inclination_effect}. The model
has  $T_{\mathrm{max}}=7500\,\mathrm{K}$ and
$\dot{M}_{\mathrm{acc}}=10^{-7}\,\mathrm{M_{\sun}\, yr^{-1}}$).  The
figure shows that the peak (normalised) flux decreases as the
inclination angle increases, as does the EW.  The line flux and the
EWs decrease mainly because the fraction of the accretion stream
blocked by the photosphere increases as the inclination increases.
The line EW in Fig.~\ref{fig:acc_inclination_effect} changes from
32 to 21\,\AA\, as the inclination changes from $10^\circ$ to
$80^\circ$, i.e. a fractional change of 0.65.  On the other hand,
the 10~per~cent width of the line changes from 290~$\kmps$ to
470~$\kmps$; hence, the fractional change is 1.6.

Because of the geometry of the
magnetospheric accretion (c.f. Fig.~\ref{fig:config_discwind}) and of
the presence of the gas with the highest velocity close to the stellar
surface, the highest red-shifted line-of-sight velocity is visible
only at the high inclination angles. This explains the wider
appearance of the profile with $i=80^{\circ}$ compared to the
relatively narrow line appearance of the profile with
$i=10^{\circ}$.  A similar EW dependence on
inclination angle is found for the models with different 
magnetospheric temperatures and the mass-accretion rates.

@@

  \caption{The dependency of the H$\alpha$ profile on the source
    displacement parameter $d$. The line profiles are computed with 
    the wind parameter $T_{\mathrm{wind}} = 9000\Kelvin$ and $\beta  =
    0.5$ for $i=55^{\circ}$.  The profiles in the figure use $d = 
   2~R_{*}$ (dashed),   $7~R_{*}$ (dash-dot) and $88~R_{*}$
    (dash-dot). All the other parameters are same as in
    Fig.~\ref{fig:atlas_discwind_only}.  As the wind becomes more
    collimated (larger $d$), the wind emission becomes stronger.  
    The P-Cygni absorption disappears for the model with a small
    collimation (e.g.~$d = 2~R_{*}$ case) since the opening angel of
    the disc wind is larger than the inclination angle. }

@@

\caption{A summary of the reference disc-wind model parameters used
in Section~\ref{sub:result-discwind-knigge}. See also
Section~\ref{sub:model-disc-wind}.}

>>>>

In the currently favoured model of accretion in CTTS, the accretion
disc is disrupted by the magnetosphere, which channels the gas from
the disc onto the stellar surface (e.g., \citealt{uchida:1985};
\citealt{koenigl:1991}; \citealt{cameron:1993};
\citealt{shu:1994}). This picture is supported by recent measurements
of strong ($\sim 10^{3}\mathrm{G}$) magnetic fields in CTTS (e.g.,
\citealt{johns-krull:1999}; \citealt{symington:2005b}) and by
radiative transfer models which reproduce the gross characteristics of
observed profiles for some CTTS (\citealt*{muzerolle:2001}). In
particular the magnetospheric accretion (MA) model explains blue-ward
asymmetric emission line profiles as resulting from the partial
occultation of the flow  by the inner part of the accretion disc and
by the stellar photosphere, while inverse 
P~Cygni profiles in the MA model result from inflowing material at
near free-fall velocities seen projected against hotspots on the
stellar surface.

@@

An additional continuum source is considered for models which include 
magnetospheric accretion: as the infalling gas approaches the stellar
surface, it decelerates in a strong shock, and is heated to $\sim
10^{6}\,\mathrm{K}$.  The X-ray radiation produced in the shock will
be absorbed by the gas locally, and re-emitted as optical and UV
(\citealt{calvet:1998}; \citealt{gullbring:2000}). This
will create hot rings where the magnetic field intersects with the
surface. We assume that the free-falling kinetic energy is thermalized
in the radiating layer, and is re-emitted as blackbody radiation with
a single temperature. With the parameters of the magnetosphere and the
star given above (Table~\ref{tab:std_parameters}), about $8$~per~cent
of the surface is covered by the hot rings. 
In our model, the fraction of hot-ring coverage area is
determined only by the geometry of the magnetosphere
(Section~\ref{sub:Magnetosphere}). \citet{gullbring:2000} estimated the
fractions of the hot-ring coverage area for three intermediate CTTS by
using the observed UV and optical spectra. They found the range of
fraction to be 0.3--3.0~per~cent.  This is smaller than the value adopted
here; however, even if we use 1~per~cent hot-ring area
coverage model (by changing the width of the magnetosphere), we
find the line shapes of H$\alpha$ do not change significantly.

If the mass-accretion rate
is $10^{-7}\,\mathrm{M_{\sun}\, yr^{-1}}$, the ratio of the accretion
luminosity to the photospheric luminosity is about $0.5$, and the
corresponding temperature of the hot rings is about
$6400\,\mathrm{K}$.  The luminosity ratio and the temperature depend
on a adopted mass-accretion rate. The continuum emission from the hot
rings is taken into account when computing the line profiles.

@@

\caption{Summary of the reference classical T~Tauri star model
  parameters. Note that a typical mass-accretion rate of CTTS is $\sim 3
  \times 10^{-8}\,\mathrm{M_{\sun}\, yr^{-1}}$
  (e.g. \citealt{gullbring:1998}; \citealt{calvet:2004}); however, a
  slightly higher mass-accretion rate is adopted here to demonstrate the 
  effect of line-broadenings more clearly (c.f. Section~\ref{sub:model-acc}). }
\label{tab:std_parameters}

@@

We adopt the MA flow model of \citet*{hartmann:1994},
as done  by \citet{muzerolle:2001} and by \citet*{symington:2005},
in which the gas accretion on to the stellar surface from the innermost
part of the accretion disc occurs through a dipolar stellar magnetic
field. The magnetic field is assumed to be so strong that the gas
flow does not affect the underlying magnetic field itself. As shown
in Fig.~\ref{fig:config_discwind} (region~2), the innermost radius ($R_{\mathrm{mi}}$)
of the magnetosphere at the equatorial plane ($z=0$) is assigned
to be same as the inner radius ($R_{\mathrm{di}}$) of the accretion
disc where the flow is truncated. In our models, $R_{\mathrm{mi}}$
and the outer radius ($R_{\mathrm{mo}}$) of the magnetosphere (at
the equatorial plane) are set to be $2.2\, R_{*}$ and $3.0\, R_{*}$
respectively. The geometry of the magnetic field/stream
lines is fixed for all calculations. We note that this magnetospheric geometry
is identical to the {}``small/wide'' model of \citet{muzerolle:2001}.

The magnetic field and the gas stream lines are assumed to have the
following simple form 
\begin{equation}
      r=R_{\mathrm{m}}\sin^{2}\theta
  \label{eq:dipole}
\end{equation}
\citep*[see][]{ghosh:1977} where $r$, and $\theta$ are the radial and
the polar component of a field point position vector in the accretion
stream.  $R_{\mathrm{m}}$ is the radial distance to the field
line at the equatorial plane ($\theta=\pi/2$), and its value is
restricted between $R_{\mathrm{mi}}$ and 
$R_{\mathrm{mo}}$ (Fig.~\ref{fig:config_discwind}). 
Using the field geometry above and conservation of
energy, the velocity and the density of the accreting gas along the
stream line are found as in \citet{hartmann:1994}.

The temperature structure of the magnetospheric used by
\citet{hartmann:1994} is adopted here. They computed the temperature,
assuming a volumetric heating rate which is proportional to $r^{-3}$,
by solving the energy balance of the radiative cooling rate of
\citet{hartmann:1982} and the heating rate \citep{hartmann:1994}.
\citet{martin:1996} presented a self-consistent determination of the
thermal structure of the inflowing gas along the dipole magnetic field
(equation~\ref{eq:dipole}) by solving the heat equation coupled to the
rate equations for hydrogen. He found that main heat source is
adiabatic compression due to the converging nature of the flow, and
the major contributors to the cooling process are bremsstrahlung
radiation and line emission from Ca~II and Mg~II ions. 
Although the model of \citet{martin:1996} seems to be more comprehensive
than that of \citet{hartmann:1994}, \citet*{muzerolle:1998} found that
the line profile models computed according to the temperature
structure of the former do not agree with observations while they do
have much better agreement when the temperature of the latter model was
used.  The temperature structure of the magnetosphere is still a big
uncertainty in the accretion model; however, for a heuristic reason,
we adopted the model of \citet{hartmann:1994}.

@@

The local mass-loss rate per unit area ($\dot{m}$) is assumed to
be proportional to the mid-plane temperature of the disc, and is a
function of the cylindrical radius $w=\left(x^{2}+y^{2}\right)^{1/2}$, i.e.
\begin{equation}
  \dot{m}\left(w\right) \propto
       \left[ T\left(w\right) \right]  ^{4\alpha}\,\,.\label{eq:discwind-massloss-temp}
\end{equation}
The mid-plane temperature of the disc is assumed to be expressed
as a power-law in $w$; thus, $T\propto w^{q}$. Using this in the
relation above, one finds 
\begin{equation}
  \dot{m}\left(w\right)\propto w^{p}
  \label{eq:discwind-massloss-w}
\end{equation}
where $p=4\alpha\times q$. The index of the mid-plane temperature power
law is adopted from the dust radiative transfer model of
\citet{whitney:2003a} who found the innermost part of the accretion
disc has $q=-1.15$.  In order to be consistent with the collimated
disc-wind model of \citet{krasnopolsky:2003} who used $p=-7/2$, the
value of $\alpha$ is set to 0.76. The constant of proportionality in
equation~\ref{eq:discwind-massloss-w} is found by integrating
$\dot{m}$ from $R_{\mathrm{wi}}$ to $R_{\mathrm{wo}}$, and 
normalising the value to the total mass-loss rate
$\dot{M}_{\mathrm{wind}}$.

The azimuthal/rotational component of the wind velocity
$v_{\phi}\left(w,z\right)$ is computed from the Keplerian rotational
velocity at the emerging point of the stream line
i.e. $v_{\phi}\left(w_{i},0\right)=\left(GM_{*}/w_{i}\right)^{1/2}$
where $w_{i}$ is the distance from the rotational axis ($z$) to the
emerging point on the disc, and by assuming the conservation of the
specific angular momentum along a stream line:
\begin{equation}
  v_{\phi}\left(w,z\right)\, = 
  v_{\phi}\left(w_{i},0\right)\,\left(\frac{w_{i}}{w}\right)\,\,.
  \label{eq:discwind-toroidal-velocity}
\end{equation}
Based on the canonical $\beta$ velocity law of hot stellar winds
(c.f.~\citealt*{castor:1975}), the poloidal component of the wind
velocity ($v_{p}$) parameterised as:
\begin{equation}
  v_{p}\left(w_{i},l\right) = c_{\mathrm{s}}\left(w_{i}\right) +
  \left[f\, v_{\mathrm{esc}}-c_{\mathrm{s}}\left(w_{i}\right)\right]\left(1-\frac{R_{s}}{l+R_{s}}\right)^{\beta}
  \label{eq:discwind-poloidal-velocity}
\end{equation}  
where $c_{\mathrm{s}}$, $f$, and $l$ are the sound speed at the wind
launching point on the disc, the constant scale factor of the
asymptotic terminal velocity to the local escape velocity (from the
wind emerging point on the disc), and the distance from the disc
surface along stream lines respectively. $R_{s}$ is the wind scale
length, and its value is set to 10~$R_{\mathrm{wi}}$ by following
\citet{long:2002}.

@@

  \caption{The dependency of the disc-wind density and velocity on the wind
    acceleration parameter $\beta$. The wind density $\rho$ (right panel)
    and the poloidal velocity component $v_{p}$ (left panel) along the
    stream line starting from the mid point of the wind launching zone,
    i.e. $\left ( w,z \right ) = \left ( w_{\mathrm{mid}}, 0 \right )$
    where $w_{\mathrm{mid}} = \left( R_{\mathrm{wi}}+R_{\mathrm{wo}}
    \right ) / 2$, are shown as a function of the distance ($l$) from the
    wind launching point (c.f.~
    equations~\ref{eq:discwind-poloidal-velocity} and
    \ref{eq:discwind-density}). The azimuthal velocity component
    ($v_{\phi}$), which is independent of $\beta$
    (c.f. equation~\ref{eq:discwind-toroidal-velocity}), is also shown in
    the right panel for comparison. The density is normalised to the
    density $\rho_{0}$ at the wind launching point for the $\beta = 1.0$ case.
    The $v_{p}$ reaches the terminal velocity by 100~au for all
    $\beta$. In the far field ($l>10$~au), the density is approximately
    proportional to $\sim l^{-2}$. }

@@

\subsubsection{Dust model}
\label{sub:dust-model}

@@

We further modified {\sc torus} by replacing the Monte Carlo line
transfer algorithm with a direct integration method
(c.f. \citealt{mihalas:1978}) for computing the observed flux as a
function of frequence.  The integration of the flux is performed in
the cylindrical coordinate system $\left(p,\, q,\, t\right)$ which is
obtained by rotating the original stellar coordinate system
$\left(\rho,\,\phi,\, z\right)$ around the $y$ axis by the inclination
of the line of sight.  Note that the $t$-axis coincides with the line
of sight with this rotation. The observed flux ($F_{\nu}$) is given
by:
\begin{equation}
  F_{\nu}=\frac{1}{4\pi d^{2}}
  \int_{0}^{p_{\mathrm{max}}}\int_{0}^{2\pi}p\,\sin q\,
  I_{\nu}\mathrm{\, d}q\,\mathrm{d}p
  \label{eq:flux_integral}
\end{equation}
where $d$, $p_{\mathrm{max}}$, and $I_{\nu}$ are the distance to
an observer, the maximum extent to the model space in the projected
(rotated) plane, and the specific intensity ($I_{\nu}$) in the direction
on observer at the outer boundary. For a given ray along $t$, the
specific intensity is given by:
\begin{equation}
       I_{\nu} = I_{0}e^{-\tau_{\infty}} + 
           \int_{t_{0}}^{t_{\infty}}S_{\nu}\left(t\right)\,
                e^{-\tau}\mathrm{d}\tau
 \label{eq:formal_sol_integral}
\end{equation}
where $I_{0}$ and $S_{\nu}$ are the intensity at the boundary on the
opposite to the observer and the source function
($\eta_{\nu}/\chi_{\nu}$) of the stellar atmosphere/wind at a
frequency $\nu$. For a ray which intersects with the stellar core,
$I_{0}$ is computed from the stellar atmosphere model of
\citet{kurucz:1979} as described in
Section~\ref{sub:Continuum-Source}, and $I_{0}=0$ otherwise. If the
the ray intersects with the hot ring on the stellar surface created by
the accretion stream, we set
$I_{0}=B_{\nu}\left(T_{\mathrm{ring}}\right)$ where $B_{\nu}$ is the
Planck function and $T_{\mathrm{ring}}$ is the temperature of the hot
ring.  The initial position of each ray is assigned to be at the
centre of the surface element ($\mathrm{d}A=p\,\sin
q\,\mathrm{d}q\,\mathrm{d}p$).  The code execution time is
proportional to $n_{p}\, n_{q}\, n_{\nu}$ where $n_{p}$ and $n_{q}$
are the number of cylindrical radial and angular points for the flux
integration, and $n_{\nu}$ is the number of frequency points. In the
models presented in the following section, $n_{p}=180$, $n_{q}=100$,
and $n_{\nu}=101$ are used unless specified otherwise. A linearly
spaced radial grid is used for the area where the ray intersects with
magnetosphere, and a logarithmically spaced grid is used for the wind
and the accretion disc regions.

@@

For a point in the magnetosphere and the wind flows, the emissivity
and the opacity of the media are given as:
\begin{equation}
  \left\{ 
  \begin{array}{rcl}
    \eta_{\nu} & = & \eta_{\mathrm{c}}^{\mathrm{H}}+\eta_{l}^{\mathrm{H}}\\
    \chi_{\nu} & = & \chi_{c}^{\mathrm{H}}+\chi_{l}^{\mathrm{H}}
    +\sigma_{\mathrm{es}}
  \end{array}
  \right.
  \label{eq:emiss_opa_gas}
\end{equation}
where $\eta_{\mathrm{c}}^{\mathrm{H}}$ and $\eta_{l}^{\mathrm{H}}$ are
the continuum and line emissivity of
hydrogen. $\chi_{c}^{\mathrm{H}}$, $\chi_{l}^{\mathrm{H}}$, and
$\sigma_{\mathrm{es}}$ are the continuum, line opacity
(equation~\ref{eq:line_opacity}) of hydrogen, and the electron
scattering opacity. Similarly, for a point in the accretion disc,
\begin{equation}
  \left\{ 
  \begin{array}{rcl}
    \eta_{\nu} & = & 0\\
    \chi_{\nu} & = & \kappa_{\mathrm{abs}}^{\mathrm{dust}}
    +\kappa_{\mathrm{sca}}^{\mathrm{dust}}
  \end{array}
  \right.
  \label{eq:emiss_opa_dust}
\end{equation}
where $\kappa_{\mathrm{abs}}^{\mathrm{dust}}$ and
$\kappa_{\mathrm{sca}}^{\mathrm{dust}}$ are the dust absorption, and
scattering opacity which are calculated using the dust property
described in Section~\ref{sub:dust-model}.  We neglect the
emissivity from the disc: since the disc masses of CTTS are rather small
($\sim 0.01\,\mathrm{M_{\sun}}$) and low temperature ($\lesssim 
1600\,\mathrm{K}$), the continuum flux contribution at H$\alpha$
wavelength is expected to be negligible (e.g.~\citealt{chiang:1997}).

@@

Using the reference model parameters (Table~\ref{tab:std_parameters})
for the central star and the magnetosphere, we examine the dependency
of H$\alpha$ on the temperature ($T_{\mathrm{max}}$) of the accretion
flow and the mass accretion rate ($\dot{M}_{\mathrm{acc}}$), as done
by \citet{muzerolle:2001} for H$\beta$. 
The temperature structure of \citet{hartmann:1994} is  adopted here,
and the parameter $T_{\mathrm{max}}$ sets the maximum temperature of the accretion
streams. 
We note that here we have
computed the hot ring temperature self-consistently, whereas
\citet{muzerolle:2001} used a constant hot ring temperature
($8000\,\mathrm{K}$) for most of their models.  The accretion
luminosity ($L_{\mathrm{acc}}$) for models with
$\dot{M}_{\mathrm{acc}}=10^{-7}\,\mathrm{M_{\sun}\, yr^{-1}}$ is about
a half of the photospheric luminosity.

@@

The dependency of the line profile on inclination angle ($i$) is
demonstrated in Fig.~\ref{fig:acc_inclination_effect}. The model
has  $T_{\mathrm{max}}=7500\,\mathrm{K}$ and
$\dot{M}_{\mathrm{acc}}=10^{-7}\,\mathrm{M_{\sun}\, yr^{-1}}$).  The
figure shows that the peak (normalised) flux decreases as the
inclination angle increases, as does the EW.  
The line flux and the EWs decrease mainly because (1)~the fraction of
the accretion stream blocked by the photosphere increases as the
inclination increases (c.f.~Fig.~\ref{fig:config_discwind}, lower-left
part of the accretion stream i.e.~$z<0$ and $x<0$), and (2)~the
fraction of the accretion stream blocked by the inner part of the
accretion disc increases as the inclination increases
(c.f.~Fig.~\ref{fig:config_discwind}, lower-right part of accretion
stream i.e.~$z<0$ and $x>0$). While the photosphere mostly blocks the
blue-shifted part of the stream, the inner disc blocks the red-shifted
part.
The line EW in Fig.~\ref{fig:acc_inclination_effect} changes from
32 to 21\,\AA\, as the inclination changes from $10^\circ$ to
$80^\circ$, i.e.~a fractional change of 0.65.  On the other hand,
the 10~per~cent width of the line changes from 290~$\kmps$ to
470~$\kmps$; hence, the fractional change is 1.6.  
A similar EW dependence on
inclination angle is found for the models with different 
magnetospheric temperatures and the mass-accretion rates.

Because of the geometry of the
magnetospheric accretion (c.f. Fig.~\ref{fig:config_discwind}) and of
the presence of the gas with the highest velocity close to the stellar
surface, the highest blue-shifted line-of-sight velocity is seen 
only at the high inclination angles. For example, the
accretion stream close to the photosphere ($\sim 1R_{*}$ above the
photosphere) on the upper-left
quadrant in Fig.~\ref{fig:config_discwind} ($x<0$ and $z>0$) will be
more aligned with the line-of-sight of an observer in the upper-right
quadrant in the same figure ($x>0$ and $z>0$) as the inclination
increases. This results in a slight increase in the blue wing flux for
a higher inclination model as seen in
Fig.~\ref{fig:acc_inclination_effect}.

@@

  \caption{The dependency of the H$\alpha$ profile on the source
    displacement parameter $d$. The line profiles are computed with 
    the wind parameter $T_{\mathrm{wind}} = 9000\Kelvin$ and $\beta  =
    0.5$ for $i=55^{\circ}$.  The profiles in the figure use $d = 
   2~R_{*}$ (dashed),   $7~R_{*}$ (dash-dot) and $22~R_{*}$
    (solid). All the other parameters are same as in
    Fig.~\ref{fig:atlas_discwind_only}.  As the wind becomes more
    collimated (larger $d$), the wind emission becomes stronger.  
    The P-Cygni absorption disappears for the model with a small
    collimation (e.g.~$d = 2~R_{*}$ case) since the opening angle of
    the disc wind is larger than the inclination angle. }

@@

\caption{A summary of the reference disc-wind model parameters used
in Sections~\ref{sub:result-discwind-knigge} and \ref{sub:result-discwind-hybrid}.
See also Section~\ref{sub:model-disc-wind}.}
